Racial and Ethnic Differences in ETI

### 2023-09-29 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
mutations in eligibility (need to update survival)

```{r data cleaning}
#| include: false
# source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
load("UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
```

@@

```{r results='asis'}
set_gtsummary_theme(theme_gtsummary_compact(), quiet = T)
years <- split.data.frame(annual, annual$ReviewYear)
invisible(lapply(names(years), function(n) {
  cat("\n")
  cat("###", n)
  cat("\n")
  d <- years[[n]]
  t <- d %>%
    select(all_of(t1_vars), race) %>%
    tbl_summary(by = race, missing_text = "Missing") %>%
    bold_labels()
  print(t)
  cat("\n")
}))
reset_gtsummary_theme()
```

@@

## Survival analysis

For the following analysis, the start time ($t=0$) was the approximate date that each person became eligible for ETI. Because we don't have exact dates of birth, I calculated an approximate DOB based on age and encounter dates. Encounters prior to ETI eligibility were excluded, and only time to first encounter on ETI was considered (i.e. we did not fit a multi-state model to account for those who started and then stopped the medication).

```{r}
# Create survival dataset
surv_dat <- encounter %>%
  mutate(stop = as.numeric(encounterdate - eti_elig_date)) %>%
  filter(stop >= 0) %>%
  group_by(eDWID) %>%
  mutate(start = lag(stop, default = 0)) %>%
  select(eDWID, race, encounterdate, encounterage, eti_elig_date, Vx445comb, start, stop) %>%
  filter(stop <= first(stop[Vx445comb == "Yes"]))
surv_dat$Vx445comb <- as.numeric(factor(surv_dat$Vx445comb, levels = c("No", "Yes"))) - 1
# Fit
fit <- survfit(Surv(start, stop, Vx445comb, type = "interval") ~ race,
  id = eDWID, data = surv_dat
)
# Plot
names(fit$strata) <- gsub("race=", "", names(fit$strata))
fit %>%
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  add_risktable(
    risktable_stats =
      c("{round(n.risk)}", "{round(cum.event)}"),
    stats_label = c("Eligible", "Num. Started")
  ) +
  add_risktable_strata_symbol() +
  add_quantile() +
  ylab("Proportion on Vx445comb") + xlab("Days From Eligibility")
```

Without Asian subgroup for clearer plot:

```{r}
fit <- survfit(Surv(start, stop, Vx445comb, type = "interval") ~ race,
  id = eDWID, data = surv_dat[surv_dat$race != "Asian", ]
)
# Plot
names(fit$strata) <- gsub("race=", "", names(fit$strata))
fit %>%
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  add_risktable(
    risktable_stats =
      c("{round(n.risk)}", "{round(cum.event)}"),
    stats_label = c("Eligible", "Num. Started")
  ) +
  add_risktable_strata_symbol() +
  add_quantile() +
  ylab("Proportion on Vx445comb") + xlab("Days From Eligibility")
```

>>>>

```{r data cleaning}
#| include: false
# source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
load("/Users/timvigers/Dropbox/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
```

@@

```{r results='asis'}
#| message: false
set_gtsummary_theme(theme_gtsummary_compact(), quiet = T)
years <- split.data.frame(annual, annual$ReviewYear)
invisible(lapply(names(years), function(n) {
  cat("\n")
  cat("###", n)
  cat("\n")
  d <- years[[n]]
  t <- d %>%
    select(all_of(t1_vars), race) %>%
    tbl_summary(by = race, missing_text = "Missing") %>%
    bold_labels()
  print(t)
  cat("\n")
}))
reset_gtsummary_theme()
```

@@

<!-- ## Survival analysis -->

<!-- For the following analysis, the start time ($t=0$) was the approximate date that each person became eligible for ETI. Because we don't have exact dates of birth, I calculated an approximate DOB based on age and encounter dates. Encounters prior to ETI eligibility were excluded, and only time to first encounter on ETI was considered (i.e. we did not fit a multi-state model to account for those who started and then stopped the medication). -->

<!-- ```{r} -->
<!-- # Create survival dataset -->
<!-- surv_dat <- encounter %>% -->
<!--   mutate(stop = as.numeric(encounterdate - eti_elig_date)) %>% -->
<!--   filter(stop >= 0) %>% -->
<!--   group_by(eDWID) %>% -->
<!--   mutate(start = lag(stop, default = 0)) %>% -->
<!--   select(eDWID, race, encounterdate, encounterage, eti_elig_date, Vx445comb, start, stop) %>% -->
<!--   filter(stop <= first(stop[Vx445comb == "Yes"])) -->
<!-- surv_dat$Vx445comb <- as.numeric(factor(surv_dat$Vx445comb, levels = c("No", "Yes"))) - 1 -->
<!-- # Fit -->
<!-- fit <- survfit(Surv(start, stop, Vx445comb, type = "interval") ~ race, -->
<!--   id = eDWID, data = surv_dat -->
<!-- ) -->
<!-- # Plot -->
<!-- names(fit$strata) <- gsub("race=", "", names(fit$strata)) -->
<!-- fit %>% -->
<!--   ggsurvfit(type = "risk") + -->
<!--   add_confidence_interval() + -->
<!--   add_risktable( -->
<!--     risktable_stats = -->
<!--       c("{round(n.risk)}", "{round(cum.event)}"), -->
<!--     stats_label = c("Eligible", "Num. Started") -->
<!--   ) + -->
<!--   add_risktable_strata_symbol() + -->
<!--   add_quantile() + -->
<!--   ylab("Proportion on Vx445comb") + xlab("Days From Eligibility") -->
<!-- ``` -->

<!-- Without Asian subgroup for clearer plot: -->

<!-- ```{r} -->
<!-- fit <- survfit(Surv(start, stop, Vx445comb, type = "interval") ~ race, -->
<!--   id = eDWID, data = surv_dat[surv_dat$race != "Asian", ] -->
<!-- ) -->
<!-- # Plot -->
<!-- names(fit$strata) <- gsub("race=", "", names(fit$strata)) -->
<!-- fit %>% -->
<!--   ggsurvfit(type = "risk") + -->
<!--   add_confidence_interval() + -->
<!--   add_risktable( -->
<!--     risktable_stats = -->
<!--       c("{round(n.risk)}", "{round(cum.event)}"), -->
<!--     stats_label = c("Eligible", "Num. Started") -->
<!--   ) + -->
<!--   add_risktable_strata_symbol() + -->
<!--   add_quantile() + -->
<!--   ylab("Proportion on Vx445comb") + xlab("Days From Eligibility") -->
<!-- ``` -->
